---
title: "Ebola Superspreading Analysis"
author: "Oluwatobi Olatunbosun"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 1. Load Libraries

```{r load-libraries}
library(epicontacts)
library(fitdistrplus)
library(superspreading)
library(tidyverse)
library(outbreaks)
```

## 2. Load and Prepare Ebola Data

```{r load-data}
# Retrieve Ebola simulated line list and contacts data
ebola_linelist <- outbreaks::ebola_sim_clean$linelist
ebola_contact <- outbreaks::ebola_sim_clean$contacts

# Create an epicontacts object
ebola_epicontacts <- epicontacts::make_epicontacts(
  linelist = ebola_linelist,
  contacts = ebola_contact,
  directed = TRUE
)
```

## 3. Visualize the Contact Network

```{r plot-epicontacts}
# Plot the contact network
plot(ebola_epicontacts)
```

## 4. Analyze Secondary Cases (Out-Degree)

```{r secondary-cases}
# Get secondary case counts (number of people each case infected)
secondary_cases <- epicontacts::get_degree(
  x = ebola_epicontacts,
  type = "out",
  only_linelist = TRUE
)

# Print secondary cases
secondary_cases
```

## 5. Plot Degree Distribution

```{r plot-distribution}
# Histogram of secondary cases
hist(secondary_cases, 
     breaks = 30,
     main = "Degree Distribution of Secondary Cases",
     xlab = "Number of Secondary Cases",
     ylab = "Frequency",
     col = "lightblue",
     border = "black")
```

## 6. Fit a Negative Binomial Distribution

```{r fit-nbinom}
# Fit negative binomial to secondary cases
fitnb <- fitdistrplus::fitdist(secondary_cases, "nbinom")
print(fitnb)
```

> The negative binomial distribution is characterized by two
> parameters: - `size` (k): Dispersion parameter - `mu`: Mean number of
> secondary cases
>
> Example estimates might look like: - size = 0.854 - mu = 0.368

## 7. Superspreading Risk Assessment

```{r superspreading}
# Estimate probability of clusters from backward tracing
superspreading::proportion_cluster_size(
  R = fitnb$estimate["mu"],
  k = fitnb$estimate["size"],
  cluster_size = c(5, 10, 15)
)
```

## 8. Interpretation

> Using the estimated parameters (R ≈ 0.368, k ≈ 0.854), the chance of a
> backward-traced case causing: - ≥5 secondary cases: \~2.6% - ≥10 or
> ≥15 secondary cases: near 0%
>
> ### 📌 Implication:
>
> Backward tracing likely provides limited added value for containing
> Ebola transmission in this scenario, as the superspreading potential
> is low.
